---
title: "FRUBASE"
output: html_notebook
---

July 2007. Freiburg. Nov 2007, Sevilla.
Revision for v. 5 started in Sevilla, 17 Jan 2015.

### Data input, etc.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r dataset}
# Get the data from my GitHub repository.
require(downloader)
link = "https://raw.githubusercontent.com/pedroj/frubase/master/frubase.txt"
file = "frubase.txt"
if(!file.exists(file)) download(link, file, mode = "wb")
frubase <- read.table(file, sep = "\t", dec = ".", 
                      header = TRUE, na.strings="NA")

```

Prepare dataset.

```{r data_edit, echo=TRUE}
# Add column with species names to use in taxa checks.
species<- paste(frubase$gen,frubase$sp,sep=" ")
frubase<- data.frame(frubase[,1:6], species, frubase[,7:32])
# str(frubase)

```


### Filtering 

```{r filter, echo=TRUE}
require(dplyr)
frubase_df<- tbl_df(frubase)
head(frubase_df)
glimpse(frubase_df)

frubase_df %>%
        filter(gen== "Vaccinium") 

filter(frubase_df, diam > 12.0)[,7]

by_fam.diam <- group_by(frubase_df, fam)
summarise(by_fam.diam,
                count = n(),
                diam.avg = mean(diam, na.rm = TRUE),
                diam.se = mean(diam, na.rm = TRUE))
```

### Taxa names check
#### Cleanup list 

There are many ways to resolve taxonomic names in `taxize`. Of course, the ideal name resolver will do the work behind the scenes for you so that you dont have to do things like fuzzy matching. There are a few services in taxize like this we can choose from: One is the Taxonomic Name Resolution Service from iPlant (see function `tnrs`).

```{r cleanup, echo=TRUE, cache=TRUE}
# The tnrs function accepts a vector of 1 or more
mylist_tnrs <- tnrs(query=splist, getpost="POST", source = "iPlant_TNRS")

# Remove some fields
mylist_tnrs <- mylist_tnrs[,!names(mylist_tnrs) %in%
               c("matchedName","annotations","uri")]
mylist_tnrs_df<- tbl_df(mylist_tnrs)

mylist_final <- as.character(mylist_tnrs$acceptedname)

```

Note the scores. They suggest that there were no perfect matches, but they were all very close, ranging from 0.77 to 0.99 (1 is the highest). 
Let's assume the names in the "acceptedName" column are correct (and they should be).

Here is code to check a specific database:

```{r select_source}
sources <- gnr_datasources()
eol <- sources$id[sources$title == 'EOL'] #c('ncbi','itis','col','tropicos')
out <- gnr_resolve(splist, data_source_ids=eol, stripauthority=TRUE)
unique(out$results)

```

### Datasets and analyses 
```{r analyses, echo=TRUE}
nfrubase<-frubase[,18:32]             # Numeric data only

#-------------------------------------------------------------------------
### Correlation matrices
frubcor <- cor(log(nfrubase)+1,use="pairwise") # Full correlation matrix
frubmorphcor<- cor(log(nfrubase[,1:8])+1,use="pairwise") # Morph correlation matrix
frubnutrcor<- cor(log(nfrubase[,9:15])+1,use="pairwise") # Nutrient correlation matrix

#-------------------------------------------------------------------------
### Plots of correlation matrices
library(Hmisc)
round(frubcor <- cor(log(nfrubase)+1, use="pairwise"), 3)  # NOTE: pairwise
round(frubmorphcor <- cor(log(nfrubase[,1:8])+1, use="pairwise"), 3)  # NOTE: pairwise
round(frubnutrcor <- cor(log(nfrubase[,9:15])+1, use="pairwise"), 3)  # NOTE: pairwise
### the heat maps of the correlation matrix
heatmap(frubcor,symm = TRUE,margin=c(6,6),col=gray((32:0)/32))# with reorder()
heatmap(frubmorphcor,symm = TRUE,margin=c(6,6),col=gray((32:0)/32))# with reorder()
heatmap(frubnutrcor,symm = TRUE,margin=c(6,6),col=gray((32:0)/32))# with reorder()

#---------------------------------------------------------------------------
# Integration --- MORPHOLOGICAL MATRIX
	feigen_morph<-eigen(frubmorphcor)
	feigen_morph$values
fintegr.morph<-var(feigen_morph$values)
fintegr.morph
# Relative integration
fmm1<-(ncol(frubmorphcor)-1)/length(nfrubase[,1]) # Maximum expected variance
fintegr.morph/ncol(frubmorphcor)

#---------------------------------------------------------------------------
# Integration --- NUTRIENT MATRIX
	feigen_nutr<-eigen(frubnutrcor)
	feigen_nutr$values
fintegr.nutr<-var(feigen_nutr$values)
fintegr.nutr
# Relative integration
fmm2<-(ncol(frubnutrcor)-1)/length(nfrubase[,9]) # Maximum expected variance
fintegr.nutr/ncol(frubnutrcor)

#---------------------------------------------------------------------------
### Phylogenetic analyses
names(frubase)
names(nfrubase)
hist(log(nfrubase[,1]+1))
correl.frub<-correlogram.formula(log(frubase[,19])+1~cl/scl/ord/fam/gen,use="pairwise.complete.obs")
#---------------------------------------------------------------------------

```






