---
title: "Exploring the FRUBASE fleshy-fruit trait database"
author: "Pedro Jordano"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Exploring the FRUBASE fleshy-fruit trait database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4.5,
  fig.align = "center"
)
# The vignette degrades gracefully if a suggested package is absent.
has_dplyr   <- requireNamespace("dplyr", quietly = TRUE)
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
has_knitr   <- requireNamespace("knitr", quietly = TRUE)
```

## Introduction

`frubase` ships the **FRUBASE** database: fleshy-fruit traits for over 900
Angiosperm plant species worldwide, compiled from published literature and
original fieldwork (Jordano, P. 1995. *American Naturalist* **145**: 163-191).
Each record combines taxonomy, fruit morphology, pulp and seed nutritional
composition, disperser type, and geographic origin.

Loading the data is a single call  -  no download step is required, the dataset
travels with the package:

```{r load}
library(frubase)
data(frubase)
dim(frubase)
```

The 32 columns fall into four groups: taxonomy (`cl` ... `sp`, plus label and
code columns), a disperser classification (`dispcat`, `disp`), the geographic
`area`, and the numeric trait block (`leng` ... `fib`).

```{r glimpse}
str(frubase[, c("fam", "gen", "sp", "dispcat", "area",
                "leng", "diam", "frfm", "lip", "pro")])
```

## Composition of the database

How many species fall into each disperser category, and where do they come
from?

```{r counts, eval = has_dplyr}
library(dplyr)

frubase |>
  count(dispcat, sort = TRUE)

frubase |>
  count(area, sort = TRUE) |>
  head(8)
```

## Filtering by taxon

The taxonomic columns make it straightforward to pull all records for a genus:

```{r filter, eval = has_dplyr}
frubase |>
  filter(gen == "Prunus") |>
  select(gen, sp, fam, dispcat, area, leng, diam, frfm)
```

## Fruit size by plant family

A common summary is mean fruit diameter per family. This reproduces the
family-level aggregation from the original FRUBASE analysis workflow:

```{r family-summary, eval = has_dplyr}
by_fam <- frubase |>
  group_by(fam) |>
  summarise(
    n        = n(),
    diam_avg = mean(diam, na.rm = TRUE),
    diam_se  = sd(diam, na.rm = TRUE) / sqrt(sum(!is.na(diam))),
    .groups  = "drop"
  ) |>
  filter(n >= 5) |>
  arrange(desc(diam_avg))

if (has_knitr) {
  knitr::kable(head(by_fam, 10), digits = 2,
               caption = "Ten families with the largest mean fruit diameter (n >= 5).")
} else {
  head(by_fam, 10)
}
```

## Fruit size across disperser types

Do bird-, mammal-, and mixed-dispersed fruits differ in size and pulp
chemistry?

```{r disperser-summary, eval = has_dplyr}
frubase |>
  filter(dispcat %in% c("Birds", "Mammals", "Mixed")) |>
  group_by(dispcat) |>
  summarise(
    n         = n(),
    mean_len  = mean(leng, na.rm = TRUE),
    mean_diam = mean(diam, na.rm = TRUE),
    mean_lip  = mean(lip,  na.rm = TRUE),
    mean_pro  = mean(pro,  na.rm = TRUE),
    .groups   = "drop"
  )
```

```{r size-plot, eval = has_dplyr && has_ggplot2, fig.cap = "Fruit length vs. diameter, coloured by disperser type."}
library(ggplot2)

frubase |>
  filter(!is.na(leng), !is.na(diam),
         dispcat %in% c("Birds", "Mammals", "Mixed")) |>
  ggplot(aes(leng, diam, colour = dispcat)) +
  geom_point(alpha = 0.4, size = 1.6) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_colour_manual(values = c(Birds   = "#1b7837",
                                 Mammals = "#762a83",
                                 Mixed   = "#e08214")) +
  labs(x = "Fruit length (mm)", y = "Fruit diameter (mm)",
       colour = "Disperser") +
  theme_bw()
```

## Trait correlations and integration

The numeric trait block supports morphology / nutrient correlation analyses.
Following the original workflow, correlations are computed on log-transformed
traits with pairwise-complete observations. Phenotypic *integration* is
summarised as the variance of the eigenvalues of the correlation matrix  -  a
larger value indicates traits that covary more tightly.

```{r integration}
numeric_block <- frubase[, c("diam", "frfm", "pdm", "sdm", "seeds", "seedm",
                             "ry", "kjg")]
cmat <- cor(log(numeric_block + 1), use = "pairwise.complete.obs")
round(cmat, 2)

eig <- eigen(cmat, only.values = TRUE)$values
integration <- var(eig)
cat("Eigenvalue variance (integration):", round(integration, 3), "\n")
cat("Relative integration:", round(integration / ncol(cmat), 3), "\n")
```

```{r heatmap, fig.cap = "Correlation structure among morphological traits (log scale)."}
heatmap(cmat, symm = TRUE, margins = c(6, 6),
        col = gray((32:0) / 32))
```

## Session information

```{r session}
sessionInfo()
```
