---
title: "FRUBASE"
output:
  html_document:
    theme: united
    highlight: default
---

July 2007. Freiburg. Nov 2007, Sevilla. Nov 2018, Sevilla. Revision for v. 5 started in
Sevilla, 17 Jan 2017.

### Data input, etc.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your
cursor inside it and pressing *Cmd+Shift+Enter*.

```{r dataset}
# Get the data from the GitHub repository.
require(downloader)
require(here)
link <-
"https://raw.githubusercontent.com/pedroj/FRUBASE/master/inst/extdata/frubase.tsv"
file <- "frubase.tsv"
if(!file.exists(file)) download(link, file, mode = "wb")
frubase <- read.table(file, sep = "\t", dec = ".", 
                      header = TRUE, na.strings="NA")

save(frubase, file= here::here("data/frubase.RData"))

```

Prepare dataset.

```{r data_edit, echo=TRUE}
# Add column with species names to use in taxa checks.
if (!has_name(frubase, "species")) {
  frubase= frubase |> 
    		mutate(species= paste(frubase$gen, frubase$sp,sep=" "))
} else {
  # Optionally, you might want to do something if the column already exists
  print("Column 'species' already exists.")
}

str(frubase)

```

### Filtering

```{r filter, echo=TRUE}
require(dplyr)
glimpse(frubase)

frubase |>
        dplyr::filter(gen== "Prunus") 

by_fam.diam <- frubase |> dplyr::group_by(fam) |>
dplyr::summarise(count = n(),
                diam.avg = mean(diam, na.rm = TRUE),
                diam.se = sd(diam, na.rm = TRUE) / sqrt(n()))
kable(by_fam.diam)
```

### Taxa names check

#### Cleanup list

There are many ways to resolve taxonomic names in `taxize`. Of course, the ideal name
resolver will do the work behind the scenes for you so that you don't have to do things
like fuzzy matching. There are a few services in taxize like this we can choose from: One
is the Taxonomic Name Resolution Service from iPlant (see function `tnrs`).

```{r cleanup, echo=TRUE, cache=TRUE}
library(TNRS)
# The tnrs function accepts a vector of 1 or more
splist<- frubase$species
tol_resolve(splist)

# Remove some fields
mylist_tnrs <- mylist_tnrs[,!names(mylist_tnrs) %in%
               c("matchedName","annotations","uri")]
mylist_tnrs_df<- tbl_df(mylist_tnrs)

mylist_final <- as.character(mylist_tnrs$acceptedname)

```

Note the scores. They suggest that there were no perfect matches, but they were all very
close, ranging from 0.77 to 0.99 (1 is the highest). Let's assume the names in the
"acceptedName" column are correct (and they should be).

Here is code to check a specific database:

```{r select_source}
sources <- gna_data_sources()
sources$title
eol <- sources$id[sources$title == 'Plant'] #c('ncbi','itis','col','tropicos')
out <- gnr_resolve(splist, data_source_ids=eol, stripauthority=TRUE)
unique(out$results)

```

### Datasets and analyses

```{r analyses, echo=TRUE}
nfrubase<-frubase[,18:32]             # Numeric data only

#-------------------------------------------------------------------------
### Correlation matrices
frubcor <- cor(log(nfrubase)+1,use="pairwise") # Full correlation matrix
frubmorphcor<- cor(log(nfrubase[,1:8])+1,use="pairwise") # Morph correlation matrix
frubnutrcor<- cor(log(nfrubase[,9:15])+1,use="pairwise") # Nutrient correlation matrix

#-------------------------------------------------------------------------
### Plots of correlation matrices
library(Hmisc)
round(frubcor <- cor(log(nfrubase)+1, use="pairwise"), 3)  # NOTE: pairwise
round(frubmorphcor <- cor(log(nfrubase[,1:8])+1, use="pairwise"), 3)  # NOTE: pairwise
round(frubnutrcor <- cor(log(nfrubase[,9:15])+1, use="pairwise"), 3)  # NOTE: pairwise
### the heat maps of the correlation matrix
heatmap(frubcor,symm = TRUE,margin=c(6,6),col=gray((32:0)/32))# with reorder()
heatmap(frubmorphcor,symm = TRUE,margin=c(6,6),col=gray((32:0)/32))# with reorder()
heatmap(frubnutrcor,symm = TRUE,margin=c(6,6),col=gray((32:0)/32))# with reorder()

#---------------------------------------------------------------------------
# Integration --- MORPHOLOGICAL MATRIX
	feigen_morph<-eigen(frubmorphcor)
	feigen_morph$values
fintegr.morph<-var(feigen_morph$values)
fintegr.morph
# Relative integration
fmm1<-(ncol(frubmorphcor)-1)/length(nfrubase[,1]) # Maximum expected variance
fintegr.morph/ncol(frubmorphcor)

#---------------------------------------------------------------------------
# Integration --- NUTRIENT MATRIX
	feigen_nutr<-eigen(frubnutrcor)
	feigen_nutr$values
fintegr.nutr<-var(feigen_nutr$values)
fintegr.nutr
# Relative integration
fmm2<-(ncol(frubnutrcor)-1)/length(nfrubase[,9]) # Maximum expected variance
fintegr.nutr/ncol(frubnutrcor)

#---------------------------------------------------------------------------
### Phylogenetic analyses
names(frubase)
names(nfrubase)
hist(log(nfrubase[,1]+1))
correl.frub<-correlogram.formula(log(frubase[,19])+1~cl/scl/ord/fam/gen,use="pairwise.complete.obs")
#---------------------------------------------------------------------------

```
